Summary of CA Rice Breeding

### 2017-02-07 output/summary.Rnw
Should work

\documentclass{article}
\usepackage[margins=1in]{geometry}
\usepackage{Sweave}

@@

<<echo = FALSE>>=
library(rstan)
library(ggplot2)
library(lme4)
library(rstanarm)

load("../data/all_vt_weather.Rda")
##fit <- read_stan_csv(list.files("~/Dropbox/vt_results", "^samples[0-9].csv",
##                                full.names = TRUE))

## Get the number of years in the trial
n_years <- aggregate(year ~ id, data = all_vt_result,
          function(x) {
              length(unique(x))})

first_yr <- aggregate(year ~ id, data = all_vt_result,
          min)

i <- match(all_vt_result$id, first_yr$id)
all_vt_result$first_yr <- first_yr$year[i]
all_vt_result$yrs_in_trial <- all_vt_result$year - first_yr$year[i]

@@

In general, rice yields in CA have been flat over the last 20+ years,
compared to the preceeding period. Yet, annual reports from the CA
Cooperative Rice Research Foundation show that each released variety
has a yield advantage over the previous varieties, typically on the
order of 4-6\%. This raises the question of how to reconcile these
facts. Namely, there are two possibilities:

@@

<<label=plot1, fig=TRUE, echo=FALSE>>=
boxplot(yield_lb ~ year, data = all_vt_result,
        xlab = "Year", ylab = expression("Yield ( lb ac"^-1~")"))
@

  \caption{Distribution of yield over the period 1995 -- 2015.} \label{fig:plot}
\end{figure}

\section{Do newer cultivars have a yield advantage?}

Looking at the numbers, it seems that each new medium-grain cultivar
has an advantage over the previous ones, but that the advantage is not
huge. Typical improvements of 5\% are reported by the breeders at the
RES, and the data show perhaps a bit less than this. There does not
appear to be a yield advantage of M-206 of M-205. It is worth
noting that there are fewer years of data for M-209, and those years
were typically higher yielding, which might lead ot a false optimism
regarding M-209.

\begin{figure}
<<label=plot2, fig=TRUE, echo=FALSE>>=
  vars <- grepl("^M[0-9]{3}$", all_vt_result$id)
  mod_data <- all_vt_result[vars,]

  boxplot(yield_lb ~ id, data = mod_data,
          xlab = "Cultivar",
          ylab = "Yield (lbs/ac)")
@
\caption{Boxplot of yields for four common medium grain cultivars for
  the period 1995--2015.}
\end{figure}

@@

\begin{enumerate}
  \item After a cultivar has been released, the yield decreases over
    time.
  \item Each cultivar release is recovering to the old yield, not
    increasing yield potential.
\end{enumerate}

@@

Only medium-grained, released varieties were included in the analysis,
reducing the data from \Sexpr{nrow(all_vt_result)} to
\Sexpr{nrow(mod_data)} observations. Included varieties are:
\Sexpr{paste(gsub('^M', 'M-', unique(mod_data$id)), collapse = ', ')}.

There is some evidence of a yield decrease over time, though this
turns out to be a tricky thing to estimate. This is because there are
several plausible and related mechanisms at work that are potentially
confounding eachother.

@@

% ## mod <- stan_glm(yield_lb ~ factor(year) + site + factor(year):site +
% ##                     yrs_in_trial + id + id:yrs_in_trial,
% ##                 data = mod_data, prior = normal(), cores = 2L, iter = 1000)

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{Sweave}

@@

<<echo = FALSE>>=

library(rstan)
library(ggplot2)
library(lme4)
library(rstanarm)

source("../src/prep_model_data.R")
load("h1_fit.Rda")
@

@@

In general, rice yields in CA have been flat over the last 20+ years,
compared to the preceeding period. Yet, annual reports from the CA
Cooperative Rice Research Foundation show that each released variety
has a yield advantage over the previous varieties, typically on the
order of 4--6\%. This raises the question of how to reconcile these
facts. Namely, there are two possibilities:

@@

@@

\begin{enumerate}
  \item H1: After a cultivar has been released, its yield decreases over
    time.
  \item H2: Each cultivar release is recovering to the old yield benchmark, not
    increasing yield potential.
\end{enumerate}

@@

Only medium-grained, released varieties were included in the analysis,
reducing the data from \Sexpr{nrow(all_vt_result)} to
\Sexpr{nrow(mod_data)} observations. Included varieties are:
\Sexpr{paste(gsub('^M', 'M-', unique(mod_data[,'id'])), collapse = ', ')}.

\subsection{Results}

<<fig=TRUE, echo=FALSE>>=
plot(mod, regex_pars = "yrs_in_trial")
@ 
\section{H2: Do newer cultivars have a yield advantage?}

In the above model, the estimate for $\beta_{cultivar}$ is
the estimated advantage for cultivar X at year 0 (first year in the
trial) averaged over variability attributable to differences in years,
sites, and the year:site interaction.

% Notes: the expected result here is that there is no difference
% between varieties in year 0. This might be complicated by the fact
% that we have M-202 15 years into its release. We might need to get a
% better way to cross reference these to the true start year.

\subsection{Results}

The raw numbers generally support the conclusion that each new
medium-grain cultivar has an advantage over the previous ones, but
that advantage is not large (Fig. \ref{fig:raw_yield}). Typical
improvements of 5\% are reported by the breeders at the RES, and the
data do not support gains this large.

<<label=plot1, fig=TRUE, echo=FALSE>>=
boxplot(yield_lb ~ year, data = mod_data,
        xlab = "Year", ylab = expression("Yield ( lb ac"^-1~")"))
@
  \caption{Distribution of yield over the period 1995 -- 2015.} \label{fig:plot}
\end{figure}

\begin{figure}
<<label=plot2, fig=TRUE, echo=FALSE>>=

  boxplot(yield_lb ~ id, data = mod_data,
          xlab = "Cultivar",
          ylab = "Yield (lbs/ac)")
@
\caption{Boxplot of yields for released medium grain cultivars for
  the period 1995--2015.}
\label{fig:raw_yield}
\end{figure}

@@
